\bf Title here

### 2020-06-23 paper/paper.Rmd
add a section for more general introduction

*In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints. The problem consists on optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function.... Gradient based optimization has been typically used to solve such problems. However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem....*

[Intro to optimization to set the scene for the paper? ppp]

## Derivative free optimisation

Given an objective function $f$, one way of optimising it is to equate its gradient to zero. In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and Derivative-Free Optimisation (DFO) methods can be useful to approach these problems. One common class of methods in DFO is *Direcct-search methods*. Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and has gained its popularity through its simplicity in use and reliability in solving complicated practical problems. Depends on whether a random sample is used in the search, this class of methods can be further classified as *stochastic* or *deterministic*.  The stochastic version of the direct-search methods will be the main optimisation procedure analysed in this paper.

@@

section 1: DFO

section 2: idea of diagnostic system & global object

section 3: general intro to tourr

section 4: apply diagnostic system + simulated data

section 5: plots

## Motivation to visaul diagnostics

@@

# Visual diagnostic system

@@

# Projection pursuit guided tour

@@

# Visual diagnostic in guided tour

@@

# Diagnostics plots

>>>>

In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints. The problem consists on optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function.... Gradient based optimization has been typically used to solve such problems. However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem.

Derivative free methods provides another option to optimise the objective function without evaluating any gradient or hessian information and a parciular class of methods: direct search, has gained its popularity through its conceptual simplicity.  However, the whole searching process in the algorithm remains a black-box. Plots are usually used to evaluate and compare the performance of different algorithms but it can easily become tedious because the code will have to be modified significantly when comparing different parameters in the algorithms. For example, a categorical variable with 5 levels can be easily mapped onto color while mapping another categorical variable with 30 levels will not make the plot informative. Thus the plot needs to be re-designed to better suits the characteristics of the parameter (whether the parameter is a scalar or a matrix? whether the parameter is quantitative or categorical? If categorical, how many levels does the parameter have?). This motivates the  design of a visual diagnostic framework for optimisation algorithms based on the idea of a *global object*.

The paper is organised as follows. Section \ref{DFO} gives a general literature review of optimisation, specifically derivative free optimisation. Section \ref{vis-diag} presents the new idea of constructing a
systematic visual framework that diagnoses the components of an optimisation procedure (parameters, searching path, etc). The rest of the paper serves as a comprehensive example of using the visual diagnostics on one particular problem: *projection pursuit guided tour*. Some background knowledge of projection pursuit guided tour is provided in Section \ref{tour}. Section \ref{apply} applys the concepts proposed in section \ref{vis-diag} in the tour problem and sets up the data. The last section, Section \ref{plots}, presents the visual diagnostic plots and explains how they can help to understand different aspects of the optimisation in tour.

# Derivative free optimisation {#DFO}

Given an objective function $f$, one way of optimising it is to equate its gradient to zero. In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and Derivative-Free Optimisation (DFO) methods can be useful to approach these problems. One common class of methods in DFO is *Direct-search methods*. Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and has gained its popularity through its simplicity in use and reliability in solving complicated practical problems. Depends on whether a random sample is used in the search, this class of methods can be further classified as *stochastic* or *deterministic*.  The stochastic version of the direct-search methods will be the main optimisation procedure analysed in this paper.

@@

## Difficulties

@@

# Visual diagnostic system {#vis-diag}

@@

# Projection pursuit guided tour {#tour}

@@

# Visual diagnostic in guided tour {#apply}

@@

# Diagnostics plots {#plots}

### 2020-06-23 paper/paper.tex
add a section for more general introduction

\bigskip
\begin{abstract}

\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

@@

\hypertarget{derivative-free-optimisation}{%
\subsection{Derivative free
optimisation}\label{derivative-free-optimisation}}

Given an objective function \(f\), one way of optimising it is to equate
its gradient to zero. In modern optimisation problems, gradient
information can be hard to evaluate or sometimes even impossible and
Derivative-Free Optimisation (DFO) methods can be useful to approach
these problems. The two most common methods in DFO are
\emph{direct-search methods} and \emph{model-based methods} and this
paper dedicates to the discussion of direct-search methods, which gains
its popularity due to its simplicity in use and reliability in
complicated practical problem. Coined by \citet{hooke1961direct}, direct
search methods don't require any gradient or Hessian information and
evaluate \(f\) directly. A well-known example of it is the Nelder-Mead
algorithm \citep{nelder1965simplex} and it enjoys the popularity due to
its simplicity and reliability. {[}feel like this sentence can be
expanded to include more information. xxx{]}. Further, direct search
methods can be classified as \emph{stochastic} and \emph{deterministic}
depends on whether a random sample is used in the search. The stochastic
version of direct-search method is the focus of this paper.

\hypertarget{projection-pursuit-guided-tour}{%
\subsection{Projection pursuit guided
tour}\label{projection-pursuit-guided-tour}}

@@

{[}doubt if I need the following. xxx{]}

@@

\hypertarget{problems-and-difficulties-in-pp-optimisation}{%
\subsection{problems and difficulties in PP
optimisation}\label{problems-and-difficulties-in-pp-optimisation}}

\hypertarget{visual-diagnostic-system}{%
\section{Visual diagnostic system}\label{visual-diagnostic-system}}

The idea of generalising all the diagnostic plots under one framework is
inspired by the concept of \emph{grammar of
graphic}\citep{wickham2010layered}, which powers the primary graphical
system in R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are
not defined by its appearance (i.e.~boxplot, histogram, scatter plot
etc) but by ``stacked layers''. By this design, ggplot doesn't need to
develop a gazillion of functions that each produces a different type of
plot. Instead, it creates aesthetic mapping from the variables to the
geometric objects and builds the plot through different layers.

{[}Do I need this paragraph to details the grammar of graphic? xxx{]}

@@

\hypertarget{diagnostics-plots}{%
\section{Diagnostics plots}\label{diagnostics-plots}}

@@

@@

>>>>

\bigskip
\begin{abstract}
Friedman \& Tukey commented on their intial paper on projection pursuit
in 1974 that ``the technique use for maximising the projection index
strongly influences both the statistical and the computational aspects
of the procedure.'' While many projection pursuit indices have been
proposed in the literature, few concerns the optimisation procedure. In
this paper, we developed a system of diagnostics aiming to visually
learn how the optimisation procedures find its way towards the optimum.
This diagnostic system can be applied to more general to help
practitioner to unveil the black-box in randomised iteartive
(optimisation) algorithms. An R package, ferrn, has been created to
implement this diagnostic system.
\end{abstract}

\noindent%
{\it Keywords:} optimisation, projection pursuit, guided tourr, visual, diagnostics, R
\vfill

@@

In an optimization problem the goal is to find the best solution within
the space of all feasible solutions which typically is represented by a
set of constraints. The problem consists on optimizing an objective
function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
spaced given by the problem constraints to either minimize or maximize a
function\ldots{}. Gradient based optimization has been typically used to
solve such problems. However, there are many situations where
derivatives of an objective function are unavailable or unreliable and
therefore traditional methods based on derivatives are not the best
option to solve an optimization problem.

Derivative free methods provides another option to optimise the
objective function without evaluating any gradient or hessian
information and a parciular class of methods: direct search, has gained
its popularity through its conceptual simplicity. However, the whole
searching process in the algorithm remains a black-box. Plots are
usually used to evaluate and compare the performance of different
algorithms but it can easily become tedious because the code will have
to be modified significantly when comparing different parameters in the
algorithms. For example, a categorical variable with 5 levels can be
easily mapped onto color while mapping another categorical variable with
30 levels will not make the plot informative. Thus the plot needs to be
re-designed to better suits the characteristics of the parameter
(whether the parameter is a scalar or a matrix? whether the parameter is
quantitative or categorical? If categorical, how many levels does the
parameter have?). This motivates the design of a visual diagnostic
framework for optimisation algorithms based on the idea of a
\emph{global object}.

The paper is organised as follows. Section \ref{DFO} gives a general
literature review of optimisation, specifically derivative free
optimisation. Section \ref{vis-diag} presents the new idea of
constructing a systematic visual framework that diagnoses the components
of an optimisation procedure (parameters, searching path, etc). The rest
of the paper serves as a comprehensive example of using the visual
diagnostics on one particular problem: \emph{projection pursuit guided
tour}. Some background knowledge of projection pursuit guided tour is
provided in Section \ref{tour}. Section \ref{apply} applys the concepts
proposed in section \ref{vis-diag} in the tour problem and sets up the
data. The last section, Section \ref{plots}, presents the visual
diagnostic plots and explains how they can help to understand different
aspects of the optimisation in tour.

\hypertarget{DFO}{%
\section{Derivative free optimisation}\label{DFO}}

Given an objective function \(f\), one way of optimising it is to equate
its gradient to zero. In modern optimisation problems, gradient
information can be hard to evaluate or sometimes even impossible and
Derivative-Free Optimisation (DFO) methods can be useful to approach
these problems. One common class of methods in DFO is
\emph{Direct-search methods}. Coined by \citet{hooke1961direct}, direct
search methods don't require any gradient or Hessian information and has
gained its popularity through its simplicity in use and reliability in
solving complicated practical problems. Depends on whether a random
sample is used in the search, this class of methods can be further
classified as \emph{stochastic} or \emph{deterministic}. The stochastic
version of the direct-search methods will be the main optimisation
procedure analysed in this paper.

{[}How about adding more details into derivative free methods? ppp{]}

\hypertarget{difficulties}{%
\subsection{Difficulties}\label{difficulties}}

{[}Are we using projection pursuit/guided tour to better understand the
convergency of optimization algorithms visually in combination with the
algorithms discussed below? Or we are focusing on the optimisation
problem only within the project pursuit context? Some of the problems
listed below are also applicable to optimization problem in general too.
ppp{]}

\hypertarget{vis-diag}{%
\section{Visual diagnostic system}\label{vis-diag}}

{[}I would expland this section more as the core contribution. ppp{]}

The idea of generalising all the diagnostic plots under one framework is
inspired by the concept of \emph{grammar of
graphic}\citep{wickham2010layered}, which powers the primary graphical
system in R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are
not defined by its appearance (i.e.~boxplot, histogram, scatter plot
etc) but by ``stacked layers''. By this design, ggplot doesn't need to
develop a gazillion of functions that each produces a different type of
plot. Instead, it aesthetically maps the variables to the geometric
objects and builds the plot through different layers.

\hypertarget{tour}{%
\section{Projection pursuit guided tour}\label{tour}}

@@

{[}doubt if I need the following. xxx - maybe an algorithmatic version
of guide tour would be helpful for non-expert to understand the
method?{]}

@@

\hypertarget{apply}{%
\section{Visual diagnostic in guided tour}\label{apply}}

\hypertarget{global-object-1}{%
\subsection{Global object}\label{global-object-1}}

@@

@@

{[}I would add the maths here to describe the simulations. ppp{]}

@@

\hypertarget{plots}{%
\section{Diagnostics plots}\label{plots}}
